\section{Portable Statistics-Format Conversion Interfaces}
\label{sec:mcv-adapter-app}

\subsection{Motivation}

OASIS corrects a canonical full-distribution view of a column. This is
straightforward when a DBMS exposes a self-contained histogram, but many
production systems do not. Instead, the optimizer's statistics object often
couples the histogram with auxiliary summaries whose probability mass must
remain consistent after any update.

\begin{itemize}
    \item \textbf{PostgreSQL}: Maintains a Most Common Values (MCV) list alongside a residual histogram that excludes MCV entries.
    \item \textbf{Oracle}: Uses top-frequency values together with height-balanced histograms.
    \item \textbf{SQL Server}: Employs a density vector together with histogram steps.
\end{itemize}

This coupling is the main portability obstacle for histogram-correction methods:
\emph{correcting the histogram alone} is often insufficient because the stored
statistics object is not a standalone histogram. OASIS addresses this with a
\emph{statistics-format conversion interface} that translates engine-specific
statistics layouts into a canonical full distribution, applies OASIS there, and
translates the corrected result back.

We instantiate the interface on PostgreSQL's MCV case because it is both
practically important and representative of tightly coupled statistics
architectures.

\subsection{Three-Phase Conversion Interface}

The interface follows three phases:

\begin{enumerate}
    \item \textbf{Reconstruction}: Convert engine-native statistics $\rightarrow$ canonical full distribution
    \item \textbf{Correction}: Apply OASIS on the canonical representation
    \item \textbf{Decomposition}: Convert corrected full distribution $\rightarrow$ engine-native statistics
\end{enumerate}

This design decouples OASIS's correction logic from database-specific storage
formats. PostgreSQL is the running example in this appendix, but the pattern is
more general.

\subsection{PostgreSQL Instantiation}

\subsubsection{Problem Formulation}

Given PostgreSQL statistics:
\begin{itemize}
    \item MCV list: $\mathcal{M} = \{(v_1, f_1), (v_2, f_2), \ldots, (v_k, f_k)\}$
    \item Residual histogram bounds: $\mathcal{H} = \{u_0, u_1, \ldots, u_m\}$ where adjacent values define $m$ equi-depth residual intervals
    \item Null fraction: $f_{\mathrm{null}}$
    \item Non-null invariant: $\sum_{i=1}^{k} f_i + p_{\mathrm{res}} = 1 - f_{\mathrm{null}}$
\end{itemize}

Construct a canonical full distribution $\mathcal{F}$ that exposes the complete
non-null column distribution to OASIS.

\subsubsection{Phase 1: Reconstruction}

\begin{algorithm}[H]
\caption{Reconstruct Canonical Full Distribution from PostgreSQL Statistics}
\label{alg:reconstruct}
\begin{algorithmic}[1]
\REQUIRE MCV list $\mathcal{M}$, residual bounds $\mathcal{H} = \{u_0, \ldots, u_m\}$, null fraction $f_{\mathrm{null}}$
\ENSURE Canonical full distribution $\mathcal{F} = (\mathcal{S}, \mathcal{R})$

\STATE Initialize $\mathcal{S} \leftarrow \emptyset$, $\mathcal{R} \leftarrow \emptyset$
\STATE $p_{\mathrm{res}} \leftarrow 1 - f_{\mathrm{null}} - \sum_{(v_i,f_i) \in \mathcal{M}} f_i$

\FOR{each $(v_i, f_i) \in \mathcal{M}$}
    \STATE Add singleton bucket $(v_i, f_i)$ to $\mathcal{S}$
\ENDFOR

\FOR{$j = 1$ \textbf{to} $m$}
    \STATE Add range bucket $(u_{j-1}, u_j, p_{\mathrm{res}} / m)$ to $\mathcal{R}$
\ENDFOR

\STATE Validate: $\sum_{s \in \mathcal{S}} s.\mathrm{freq} + \sum_{r \in \mathcal{R}} r.\mathrm{freq} = 1 - f_{\mathrm{null}}$
\RETURN $\mathcal{F} = (\mathcal{S}, \mathcal{R})$
\end{algorithmic}
\end{algorithm}

PostgreSQL stores histogram \emph{boundary values}, not explicit per-bucket
frequencies. The conversion interface therefore interprets the residual
histogram as $m$ equal-depth intervals carrying equal shares of the residual
non-MCV mass. Before tensorization, the interface evaluates the CDF of
$\mathcal{F}$ at OASIS's fixed quantile levels, yielding the canonical quantile
vector consumed by the model.

\subsubsection{Phase 2: OASIS Correction}

Once reconstructed, the canonical full distribution is passed unchanged to the
standard OASIS correction pipeline. The important systems property is that the
model itself remains ignorant of the source database format; all engine-specific
handling is localized to the conversion interface.

\subsubsection{Phase 3: Decomposition}

Given a corrected full distribution $\mathcal{F}' = (\mathcal{S}', \mathcal{R}')$,
reconstruct PostgreSQL-compatible statistics:
\begin{itemize}
    \item new MCV list: $\mathcal{M}'$
    \item new residual histogram bounds: $\mathcal{H}'$
    \item invariant: $\sum_i f_i' + p'_{\mathrm{res}} = 1 - f_{\mathrm{null}}$
\end{itemize}

\begin{algorithm}[H]
\caption{Decompose Canonical Full Distribution to PostgreSQL Statistics}
\label{alg:decompose}
\begin{algorithmic}[1]
\REQUIRE Full distribution $\mathcal{F}' = (\mathcal{S}', \mathcal{R}')$, MCV threshold $\tau$, MCV limit $K$, target histogram size $m{+}1$
\ENSURE MCV list $\mathcal{M}'$, residual histogram bounds $\mathcal{H}'$

\STATE \textbf{Step 1: Extract MCV candidates}
\STATE $\mathcal{C} \leftarrow \{(v, f) \mid (v, f) \in \mathcal{S}' \text{ and } f \geq \tau\}$
\STATE Sort $\mathcal{C}$ by frequency (descending)
\STATE $\mathcal{M}' \leftarrow$ top $K$ entries from $\mathcal{C}$

\STATE \textbf{Step 2: Build residual distribution}
\STATE $V_{\text{mcv}} \leftarrow \{v \mid (v, f) \in \mathcal{M}'\}$
\STATE Construct residual mixture from all range buckets in $\mathcal{R}'$ and all singletons in $\mathcal{S}'$ whose value is not in $V_{\text{mcv}}$

\STATE \textbf{Step 3: Resample equi-depth histogram bounds}
\FOR{$j = 0$ \textbf{to} $m$}
    \STATE Set $u_j$ to the residual quantile at level $j/m$
\ENDFOR
\STATE $\mathcal{H}' \leftarrow \{u_0, \ldots, u_m\}$

\STATE Validate: $\sum_{(v,f) \in \mathcal{M}'} f + p'_{\mathrm{res}} = 1 - f_{\mathrm{null}}$
\RETURN $\mathcal{M}'$, $\mathcal{H}'$
\end{algorithmic}
\end{algorithm}

\subsubsection{MCV Selection Policy}

The MCV threshold $\tau$ and limit $K$ control the final decomposition:
\begin{itemize}
    \item \textbf{High threshold}: keeps the MCV list compact but pushes more probability mass into the residual histogram
    \item \textbf{Low threshold}: retains more explicit singleton masses as MCV entries
    \item \textbf{PostgreSQL default-like regime}: $\tau \approx 0.01$, $K = 100$
\end{itemize}

This threshold is a decomposition policy choice, not part of OASIS's core
correction model. In our validation, experiments 1--3 remove threshold effects
by using threshold-free round trips, while experiment 4 isolates the effect of
$\tau_{\text{MCV}}$ after a simulated correction.

\subsection{Why This Is a Portable Interface, Not a PostgreSQL Patch}

The PostgreSQL instantiation is only one instance of the conversion pattern.
The same interface idea extends to other engines that rely on histograms to
very different degrees:

\begin{itemize}
    \item \textbf{Self-contained histograms} (MySQL, MariaDB): direct pass-through; conversion degenerates to a near-identity mapping.
    \item \textbf{MCV-coupled statistics} (PostgreSQL, Oracle): reconstruct from explicit frequent-value summaries plus residual histograms, then decompose back.
    \item \textbf{Hybrid summaries} (SQL Server): reconstruct from histogram steps plus density summaries, apply OASIS, then recompute auxiliary summaries.
\end{itemize}

The systems contribution is therefore broader than ``support PostgreSQL''.
The conversion interface turns histogram correction into a reusable component
for heterogeneous optimizer statistics architectures.

\subsection{Consistency Guarantees}

\subsubsection{Probability Mass Conservation}

\begin{theorem}[Probability Conservation]
If the source statistics satisfy the non-null invariant
$\sum_i f_i + p_{\mathrm{res}} = 1 - f_{\mathrm{null}}$, then:
\begin{enumerate}
    \item reconstruction produces a canonical full distribution with the same non-null mass;
    \item decomposition emits engine-native statistics whose MCV mass plus residual mass again equals $1 - f_{\mathrm{null}}$.
\end{enumerate}
\end{theorem}

\begin{proof}
Reconstruction assigns all explicit MCV mass to singleton buckets and all
remaining non-null mass to residual intervals by construction. Decomposition
selects an explicit MCV set and resamples the remainder as residual histogram
bounds, so the output MCV mass and residual mass partition the same corrected
non-null mass. Therefore the invariant is preserved.
\end{proof}

\subsubsection{Approximate Idempotence}

\begin{definition}[Approximate Idempotence]
Let $(\mathcal{M}', \mathcal{H}') = \mathrm{Decompose}(\mathrm{Reconstruct}(\mathcal{M}, \mathcal{H}), \tau, K)$. The conversion interface is $\epsilon$-idempotent if the round trip preserves total non-null mass within $\epsilon$ and preserves the residual quantile function within $\epsilon$ over a fixed evaluation grid.
\end{definition}

Perfect idempotence need not hold under thresholded decomposition because:
\begin{itemize}
    \item the MCV threshold may change which values remain explicit,
    \item low-frequency singleton mass may be folded back into the residual distribution,
    \item floating-point rounding accumulates.
\end{itemize}

Our threshold-free validation shows exact total-mass preservation and
machine-precision residual-quantile fidelity for the tested PostgreSQL-style
statistics.

\subsection{Implementation Interface}

We provide a Python reference interface that exposes the conversion pattern
explicitly:

\begin{lstlisting}[language=Python, caption=Statistics-Format Conversion Interface (PostgreSQL Instance)]
class PostgreSQLStatisticsAdapter:
    def to_full_histogram(
        self,
        pg_stats: PostgreSQLStatistics
    ) -> FullHistogram:
        """Phase 1: Reconstruction"""
        pass

    def from_full_histogram(
        self,
        full_hist: FullHistogram
    ) -> PostgreSQLStatistics:
        """Phase 3: Decomposition"""
        pass

    def round_trip_test(
        self,
        pg_stats: PostgreSQLStatistics
    ) -> Tuple[bool, str]:
        """Validate round-trip fidelity"""
        pass
\end{lstlisting}

The full reference interface is available in
\texttt{appendix/mcv\_adapter\_interface.py}. This appendix is the
supplementary counterpart of the compact portability discussion in the
main text and is intended to be kept numerically consistent with
\texttt{experiments/results/mcv\_validation\_results.json}.

\subsection{Performance Considerations}

\subsubsection{Time Complexity}

\begin{itemize}
    \item \textbf{Reconstruction}: $O(k + n)$ where $k = |\mathcal{M}|$ and $n = |\mathcal{H}|$
    \item \textbf{Decomposition}: dominated by MCV candidate sorting and residual quantile sampling
    \item \textbf{Total overhead}: negligible relative to correction and well within planning-time constraints in our PostgreSQL validation
\end{itemize}

\subsubsection{Space Complexity}

The canonical full distribution is temporary and small:
\[
\mathrm{Space}(\mathcal{F}) = \mathrm{Space}(\mathcal{M}) + \mathrm{Space}(\mathcal{H}) + O(1).
\]

In practice this overhead is acceptable because statistics objects are compact,
the canonical distribution exists only during correction, and no persistent
storage expansion is required.

\subsection{Validation Summary}
\label{sec:mcv-eval}

We validate the PostgreSQL instantiation of the conversion interface on
synthetic PostgreSQL-style statistics spanning uniform, normal, Zipfian, and
bimodal distributions.

\begin{table}[h]
\centering
\caption{PostgreSQL Statistics-Format Interface Validation Summary}
\label{tab:mcv-validation-summary}
\begin{tabular}{l l}
\toprule
\textbf{Metric} & \textbf{Result} \\
\midrule
Round-trip total mass error & $0$ \\
Residual quantile MAE (mean) & $2.4 \times 10^{-17}$ \\
Residual quantile MAE (max) & $9.2 \times 10^{-17}$ \\
Selectivity MAE & $1.1 \times 10^{-4}$ \\
Selectivity P99 & $3.8 \times 10^{-3}$ \\
Default-like overhead (100 MCV, 10--20 bins) & $0.066$--$0.073$ ms \\
Worst tested overhead (50 MCV, 50 bins) & $1.03$ ms \\
\bottomrule
\end{tabular}
\end{table}

These results support three claims. First, the conversion interface is exact in
mass preservation and essentially exact in residual-quantile reconstruction
under threshold-free round trips. Second, it introduces negligible estimation
bias and low latency on PostgreSQL-like settings. Third, threshold sensitivity
is graceful: increasing $\tau_{\text{MCV}}$ shrinks the MCV list and gradually
increases selectivity error, but does not compromise mass conservation.

\subsection{Takeaway}

The main value of this interface is not merely that it accommodates one
PostgreSQL-specific edge case. Rather, it makes histogram correction portable to
database systems whose optimizer statistics are stored as tightly coupled
multi-component objects. OASIS corrects the canonical full-distribution view;
the conversion interface is what makes that correction deployable across
heterogeneous DBMS statistics architectures.
