%!TEX encoding = UTF-8 Unicode
\documentclass[runningheads]{llncs}

%% ── packages ─────────────────────────────────────────────────────────────────
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\let\REQUIRE\Require
\let\ENSURE\Ensure
\let\STATE\State
\let\FOR\For
\let\ENDFOR\EndFor
\let\RETURN\Return
\usepackage{listings}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{enumitem}
\setlist{nosep,leftmargin=*}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{hyperref}

%% ── reduce spacing ───────────────────────────────────────────────────────────
\setlength{\textfloatsep}{6pt plus 2pt minus 2pt}
\setlength{\floatsep}{4pt plus 2pt minus 2pt}
\setlength{\intextsep}{4pt plus 2pt minus 2pt}
\setlength{\abovecaptionskip}{3pt}
\setlength{\belowcaptionskip}{2pt}
\setlength{\parskip}{2pt plus 1pt minus 1pt}
\setlength{\parsep}{2pt plus 1pt minus 1pt}
\setlength{\itemsep}{1pt plus 1pt minus 1pt}

%% TikZ libraries for figures
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,fit,backgrounds,calc}

%% ── listings style ───────────────────────────────────────────────────────────
\lstset{
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  frame=single,
  numbers=left, numberstyle=\tiny,
  keywordstyle=\color{blue!70!black}\bfseries,
  commentstyle=\color{gray},
  stringstyle=\color{teal},
}

\begin{document}

\title{OASIS: Supplementary Materials}

\titlerunning{OASIS: Supplementary Materials}

\author{Author Names Redacted for Review}

\authorrunning{Author Names Redacted}

\institute{Affiliation Redacted\\
\email{redacted@review.org}}

\maketitle

This document collects supplementary materials for the paper
``OASIS: Feedback-Driven Marginal Statistics Correction for Cost-Based Query Optimizers.''
It provides detailed scope and error-budget positioning (\S\,A),
including the marginal-to-joint composition boundary,
model architecture and instrumentation details (\S\,B),
baseline fairness discussion and QuickSel-H adaptation (\S\,C),
the PostgreSQL-style statistics-format adapter and conversion protocol (\S\,D),
predicate-coverage sensitivity analysis (\S\,E),
observation aggregation study details (\S\,F),
PostgreSQL planner-only statistics-injection protocol details (\S\,G),
OOD drift realism protocol (\S\,H),
benchmark-inspired DML trace protocol (\S\,I),
feedback-budget, feedback-noise, and deployment-safety diagnostic protocols
(\S\,J), and detailed overhead breakdowns (\S\,K).

%% ═════════════════════════════════════════════════════════════════════════════
\section{Scope and Error Budget}
\label{sec:scope-app}
%% ═════════════════════════════════════════════════════════════════════════════

OASIS targets \emph{single-column histogram staleness}.
We explicitly exclude from scope:
\begin{itemize}
  \item \textbf{Multi-column correlation errors}: Join cardinality misestimation
        caused by correlated column predicates requires multi-dimensional
        statistics (e.g., mvdists, DNF-based sketches, or learned joint models).
        OASIS operates on one-dimensional marginals and is complementary to
        multi-column methods.
  \item \textbf{Physical design errors}: Missing indexes, suboptimal partitioning,
        or incorrect materialized views are outside the statistics layer.
  \item \textbf{Cost-model calibration}: Even with correct selectivities, the
        optimizer's cost model may mispredict operator runtimes.
        OASIS improves selectivity estimates but does not modify the cost model.
  \item \textbf{Data distribution shifts beyond drift}: Fundamental schema changes
        (column additions, type changes) or massive bulk loads that alter the
        distribution shape beyond what a compound drift operator captures are
        out of scope; these scenarios require a fresh \texttt{ANALYZE}.
\end{itemize}

\paragraph{Error budget decomposition.}
In a production optimizer, the total cardinality estimation error decomposes as:
\[
  \text{Total Error} = \underbrace{\text{Single-Column Error}}_{\text{OASIS target}}
  + \underbrace{\text{Correlation Error}}_{\text{multi-column methods}}
  + \underbrace{\text{Cost-Model Error}}_{\text{calibration}}
  + \underbrace{\text{Physical Design Error}}_{\text{indexing/partitioning}}.
\]
OASIS addresses the first term. Even perfect single-column statistics do not
eliminate the remaining terms, but accurate marginals are a necessary
prerequisite: learned multi-column estimators (e.g., copula-based models like
CoLSE~\cite{rathuwadu2025colse}) combine one-dimensional marginals with a
learned dependence structure, making single-column statistics quality a
foundational concern.

\paragraph{Marginal-to-joint composition boundary.}
The multi-column composition experiments in the main paper should be interpreted
as downstream use of corrected marginals, not as evidence that OASIS estimates
correlations. The independence estimator uses the maximum-entropy joint
distribution implied by the supplied marginals, while the Gaussian-copula
estimator receives a fixed dependence structure. In this setting, plain OASIS
marginals improve the joint estimator only modestly and inconsistently because
they are not always feedback-consistent on the current observation window.
OASIS-Proj is therefore a deployment calibration layer: it starts from the
learned OASIS marginal correction, then applies a feedback-consistency
projection before the marginal is passed to the downstream multi-column
estimator. The core contribution remains single-column statistics correction.

%% ═════════════════════════════════════════════════════════════════════════════
%% Include the existing appendix files
%% ═════════════════════════════════════════════════════════════════════════════

\input{appendix/system_details.tex}

\section{Baseline Fairness Discussion}
\label{sec:baseline-fairness}

All feedback-driven methods compared in the main paper consume the same
observation window of size $K{=}16$ and emit corrected quantiles on the same
$B{=}10$ equi-depth histogram grid. No method receives additional information
or computational budget beyond this shared interface.

\paragraph{STHoles.}
We implement the hierarchical refinement algorithm described
in~\cite{bruno2001stholes}. STHoles builds a tree of bucket refinements from
feedback. We initialize it from the stale prior and allow up to $K$ refinement
steps per correction instance, using the same observation tuples provided to
OASIS. The output quantiles are extracted from the refined tree on the standard
$B{-}1$ grid.

\paragraph{ISOMER.}
We implement the consistency-based histogram adjustment
from~\cite{markl2007consistent}. ISOMER adjusts bucket frequencies to satisfy
observed selectivity constraints while minimizing KL divergence from the prior.
We provide the same $K{=}16$ observation tuples and extract quantiles on the
standard grid.

\paragraph{QuickSel-H.}
See \S\,\ref{sec:system-details-app} for the detailed QuickSel-H adaptation.

\paragraph{Simple heuristics.}
LinInterp and FeedAvg consume the same observation window but use fixed
aggregation rules (blend factor 0.5 and exponential moving average,
respectively). They are included to demonstrate that learned correction is
necessary, not merely that any feedback-driven approach helps.

No method receives per-table or per-distribution tuning; all use default
hyperparameters. Results are averaged over three independent training seeds
(42, 123, 456) with 95\% confidence intervals.

\input{appendix/mcv_adapter.tex}

%% ═════════════════════════════════════════════════════════════════════════════
\section{Predicate-Coverage Sensitivity Analysis}
\label{sec:coverage-sensitivity-app}
%% ═════════════════════════════════════════════════════════════════════════════

OASIS correction quality depends on the observation window covering the value
range relevant to future predicates. We conduct a controlled sensitivity
experiment at $q{=}10$ by restricting observations to a ``hot region'' of width
$w$ centered at the domain midpoint.

\begin{table}[h]
\centering
\caption{Predicate-Coverage Sensitivity at $q{=}10$}
\label{tab:coverage-sensitivity}
\begin{tabular}{c r r r}
\toprule
Coverage $w$ & Stale Prior Q-Error & OASIS Q-Error & Improvement \\
\midrule
1.00 (full) & 2.720 & \textbf{1.232} & 55\% \\
0.70        & 2.720 & \textbf{1.268} & 53\% \\
0.50        & 2.720 & \textbf{1.395} & 49\% \\
0.30        & 2.720 & \textbf{1.625} & 40\% \\
0.15        & 2.720 & \textbf{1.890} & 30\% \\
\bottomrule
\end{tabular}
\end{table}

At $w{=}0.7$ (70\% coverage), Q-Error degrades by only 3\% relative to full
coverage. At $w{=}0.3$, degradation reaches 40\% but OASIS still outperforms
the stale prior by 26\%. Even at $w{=}0.15$ (15\% coverage), OASIS provides
30\% improvement over the stale prior, demonstrating graceful degradation under
sparse feedback. The abstention policy (\S\,H in the main paper) provides a
natural guard for extreme sparsity scenarios.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Observation Aggregation Study}
\label{sec:aggregation-app}
%% ═════════════════════════════════════════════════════════════════════════════

The main paper reports that replacing attention pooling with mean/max pooling
degrades Q-Error by 12\%/18\% at $q{=}10$. Here we provide the full breakdown.

\begin{table}[h]
\centering
\caption{Observation Aggregation Ablation at $q{=}10$}
\label{tab:aggregation}
\begin{tabular}{l r r}
\toprule
Pooling Strategy & Q-Error & Degradation vs.\ Attention \\
\midrule
Attention (OASIS) & \textbf{1.232} & --- \\
Mean pooling       & 1.380 & +12\% \\
Max pooling        & 1.453 & +18\% \\
\bottomrule
\end{tabular}
\end{table}

Mean pooling treats all observations equally, diluting the signal from
highly informative predicates. Max pooling amplifies the most extreme
observation but is sensitive to outliers. Attention learns non-uniform
weights that up-weight informative observations while down-weighting
redundant or noisy ones. This advantage is most pronounced when the
observation window contains heterogeneous predicates of varying
selectivity and informativeness.

%% ═════════════════════════════════════════════════════════════════════════════
\section{PostgreSQL Planner-Only Statistics-Injection Protocol}
\label{sec:optimizer-protocol-app}
%% ═════════════════════════════════════════════════════════════════════════════

The planner-only evaluation protocol uses PostgreSQL's \texttt{EXPLAIN}
(without \texttt{ANALYZE}) to validate the
statistics$\rightarrow$estimated-cardinality$\rightarrow$plan-shape path.
\texttt{EXPLAIN} invokes PostgreSQL's optimizer and cost model, but it does not
execute the query. The protocol therefore provides optimizer-facing evidence
without making wall-clock runtime claims.

\paragraph{Protocol steps.}
\begin{enumerate}
  \item Create a synthetic PostgreSQL schema with a fact table and dimension
        table. The fact table stores the current post-drift data.
  \item Generate stale and fresh single-column distributions for
        \texttt{fact.x}; build ISOMER, OASIS, OASIS-Proj, and Hybrid corrected
        marginal statistics from the same feedback window.
  \item For each statistics state, create an analyzed one-column source table
        and copy its typed \texttt{pg\_statistic} row into the target
        \texttt{fact.x} statistics entry. This avoids hand-writing
        PostgreSQL's \texttt{anyarray} catalog fields.
  \item For each test query, obtain true cardinality using \texttt{COUNT(*)}
        and obtain PostgreSQL estimated rows and plan shape using
        \texttt{EXPLAIN (FORMAT JSON)}.
  \item Compare row-estimation Q-error, plan-shape match to the
        fresh-statistics planner, recovery of stale/fresh plan disagreements,
        and new deviations from fresh when stale already matched fresh.
\end{enumerate}

\paragraph{Results.}
The formal batch covers 12 configurations: two drift directions
(\texttt{left\_shift} and \texttt{right\_shift}), two fact-table sizes
(100K and 200K rows), and three random seeds. Each configuration evaluates
84 scan/join query templates under each statistics state. No query runtime is
measured.

\begin{itemize}
  \item Stale statistics produce aggregate row Q-error 27.748 and match the
        fresh-statistics plan shape on 56.0\% of queries.
  \item Plain OASIS reduces row Q-error to 3.306 and recovers 91.7\% of
        stale/fresh plan-shape disagreements, but introduces new plan
        deviations in 13.8\% of cases where stale statistics already matched
        fresh.
  \item OASIS-Proj reduces row Q-error to 2.360, matches the fresh-statistics
        plan shape on 96.1\% of queries, recovers 92.6\% of stale/fresh plan
        disagreements, and reduces new plan deviations to 1.1\%.
  \item Across configurations, OASIS-Proj has mean row Q-error
        $2.375\pm0.275$, close to ISOMER's $2.372\pm0.283$ and below plain
        OASIS's $3.313\pm0.228$.
\end{itemize}

This protocol validates the causal chain
$\text{statistics correction} \rightarrow \text{estimated rows}
 \rightarrow \text{cost re-estimation} \rightarrow \text{plan shape}$
inside a real PostgreSQL optimizer, while intentionally avoiding runtime
claims that would depend on executor, cache, and system effects outside the
single-column statistics layer.

\input{../experiments/results/postgres_planner_stats_injection_batch_20260529/table_postgres_plan_examples.tex}

Table~\ref{tab:postgres_plan_examples} gives representative plan-shape changes
from the batch. These examples are not used as runtime evidence. They show the
mechanism that the planner-only experiment is intended to validate: stale
statistics can move estimated rows enough to select a different root operator
or join strategy, while OASIS-Proj moves the estimated rows and plan shape back
toward the fresh-statistics planner.

%% ═════════════════════════════════════════════════════════════════════════════
\section{OOD Drift Realism Protocol}
\label{sec:ood-drift-app}
%% ═════════════════════════════════════════════════════════════════════════════

The OOD drift realism suite evaluates the same OASIS checkpoint used in the
main synthetic experiments. The checkpoint is trained only on the compound drift
generator; no samples from the OOD families are used for retraining or
hyperparameter tuning. Each OOD case follows the same causal protocol as the
main synthetic suite: capture stale quantiles, interleave drift steps with 16
feedback observations, compute final post-drift quantiles, and evaluate future
selectivity estimates against the final distribution.

The six OOD families are deployment-inspired patterns implemented in
\path{extended_drift_generators.py}: batch loading, range shift, skew
evolution, outlier burst, multimodal drift, and seasonal/mixed drift. The
reported table uses 128 held-out cases per family, $B=10$ histogram buckets,
and 64 future CDF probe points per case. This experiment is not intended to
prove that the simulator is a production DML trace; it tests whether OASIS and
OASIS-Proj remain useful when the drift mechanism differs from the training
generator.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Benchmark-Inspired DML Trace Protocol}
\label{sec:trace-drift-app}
%% ═════════════════════════════════════════════════════════════════════════════

The trace-grounded sanity check is not a production-log experiment. It
constructs explicit DML event streams whose table/column stories are anchored
to analytical benchmark maintenance patterns: append-heavy sales dates,
promotion price revisions, inventory restocking, returns and cancellations,
customer segment churn, and seasonal mixed maintenance. Values are normalized
to the unit interval so that the same single-column correction interface is
used as in the synthetic and OOD experiments.

Each trace case begins with a stale histogram, then executes 16 DML steps. Each
step emits concrete insert, update, and delete counts and then records one
predicate feedback observation against the current post-step table. The
reported run uses 96 cases per trace, $B=10$ histogram buckets, and 64 future
CDF probe points per case. The same OASIS checkpoint trained on compound drift
is used unchanged; no trace-specific retraining or hyperparameter tuning is
performed.

\input{../experiments/results/trace_grounded_drift_20260529/table_trace_grounded_drift.tex}

The table should be interpreted as a realism sanity check rather than a claim
that the traces match a particular deployment. Its main role is to expose a
useful boundary: OASIS-Proj and Hybrid consistently improve over stale
statistics, while plain OASIS can regress on mild update-heavy traces where
feedback consistency is more valuable than learned global extrapolation.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Feedback Budget, Feedback Noise, and Safety Diagnostics}
\label{sec:feedback-sensitivity-app}
%% ═════════════════════════════════════════════════════════════════════════════

The feedback-budget experiment truncates each held-out sample to the $K$ most
recent observations, with $K\in\{2,4,8,16\}$. The trained OASIS checkpoint is
not retrained for smaller budgets; instead, the truncated window is padded to
the checkpoint's native 16-slot tensor width. ISOMER, OASIS-Proj, and Hybrid
receive the same truncated observation window, so the comparison isolates the
effect of feedback availability.

The feedback-noise experiment perturbs only the observed actual selectivities
in the feedback window. It applies multiplicative log-normal noise with
$\sigma\in\{0,0.02,0.05,0.10\}$, keeps stale priors and fresh future
selectivities fixed, and evaluates the same generator-driven
optimizer-decision proxy used in the main paper. The reported diagnostic uses
64 held-out cases per drift intensity and one noise seed, because noisy
projection constraints are more expensive than the clean-feedback setting.

Both diagnostics are optimizer-facing rather than runtime-facing. The scan and
join decisions are selected from estimated selectivities under a fixed cost
proxy, then scored by true proxy cost using the generated post-drift
distribution. The purpose is to characterize deployment sensitivity, not to
claim wall-clock speedups.

The deployment-safety table in the main paper is derived from cached outputs of
four diagnostics: PostgreSQL planner-only batch summaries, feedback-budget
sensitivity, feedback-noise robustness, and the benchmark-inspired DML trace
sanity check. It does not introduce an oracle selector. Projection uses only the
current feedback window, and Hybrid chooses among stale, ISOMER, OASIS, and
OASIS-Proj by minimizing feedback-window residual. The table is generated by
\path{experiments/deployment_safety_analysis.py}.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Stage-2 Calibration: Soft Projection, Banded Projection, and the Calibrated Router}
\label{sec:stage2-calibration-app}
%% ═════════════════════════════════════════════════════════════════════════════

This section gives the protocol and full per-family results behind the main
paper's finding that the Stage-2 calibration is regime-dependent.

\subsection{Operators}

All Stage-2 operators act on the same interval partition induced by the active
feedback predicates and start from the OASIS-noProj histogram. \emph{Hard
projection} (the default OASIS Stage 2) performs cyclic I-projections so that
every active predicate's interval mass matches its observed selectivity exactly,
dropping the oldest constraints when the active set becomes infeasible.
\emph{Soft projection} instead minimizes
$\mathrm{KL}(p\,\|\,p_{\text{noProj}}) + \lambda\sum_j w_j (A_j p - y_j)^2$ by
mirror descent over cell masses, where $A_j$ is the interval mask of feedback
$j$ and $y_j$ its observed selectivity. We study three weightings: full-window
($w_j{=}1$), a fixed recent window ($w_j{=}1$ for the last $k{=}8$ observations,
$0$ otherwise), and \emph{conflict-aware}. The conflict-aware weighting fits a
hard reference distribution $p_{\text{ref}}$ to the most recent $k$ observations,
scores each constraint by $\text{conflict}_j = |A_j p_{\text{ref}} - y_j|$, and
sets $w_j = \exp(-(\text{conflict}_j/\tau)^2)$ with $\tau{=}0.03$, so feedback
contradicted by the recent observations is suppressed while old-but-consistent
feedback is retained. \emph{Banded projection} I-projects $p_{\text{noProj}}$
onto the bands $y_j\pm\kappa\sqrt{y_j(1-y_j)}$ ($\kappa{=}0$ recovers hard
projection). The \emph{calibrated router} selects, per case, the candidate with
the lowest feedback-window residual among
$\{$stale, ISOMER, OASIS-noProj, hard projection, conflict-aware soft$\}$. All
operators and the diagnostic are implemented in
\path{cdf_kll_ml_pipeline/modern_baselines.py} (\path{correct_soft_isomer},
\path{correct_band_isomer}) and \path{experiments/oasis_soft_projection_diagnostics.py}.

\subsection{Evaluation Protocol}

The single-column comparison reuses the cached compound-drift data and the fixed
OASIS checkpoint; it does not retrain. For each case it generates held-out future
predicates, estimates their selectivity from each Stage-2 output, and reports the
geometric-mean Q-error, together with the feedback residual on the observation
window, quantile MAE against fresh statistics, and an optimizer-proxy join
regret. This future-predicate aggregation differs from the
\texttt{single\_column\_projection} table in the main results (which reports a
different per-case aggregation), so the numbers here are not directly comparable
to that table and are presented as a self-contained operator comparison. The
downstream and safety suites (OOD drift, DML trace, FactorJoin, composition
family, optimizer proxy, PostgreSQL planner subset) reuse the same protocols
documented in the preceding sections, with the soft/router methods added.

\subsection{Single-Column and Optimizer-Proxy Results}

\input{../experiments/results/stage2_calibration_20260531/table_stage2_calibration_single.tex}

Table~\ref{tab:stage2-app-single} shows that full-window soft has the best raw
single-column accuracy but, as the safety suites below show, is unsafe; the
calibrated router recovers most of that accuracy (1.395) while remaining
deployment-safe, and has the lowest feedback residual, join regret, and (not
shown) new-risk loss (1.37\% versus hard 1.81\%). The router routes to the
conflict-aware soft candidate on 44\% of cases and to hard/ISOMER/OASIS-noProj
otherwise. As a negative control, banded projection monotonically \emph{worsens}
single-column Q-error as the band widens: $\kappa\in\{0.04,0.10,0.20,0.35\}$
gives Q-error $\{1.437,1.450,1.470,1.509\}$, confirming that the gain comes from
per-case routing, not from relaxing the feedback constraints.

\subsection{Out-of-Distribution and Trace Safety}

\input{../experiments/results/stage2_calibration_20260531/table_stage2_calibration_safety.tex}

Table~\ref{tab:stage2-app-safety} documents why soft projection needs a
consistency gate. Full-window soft softly fits stale, contradicted feedback and
fails on abrupt drift (batch load 1.482, range shift 1.284) and on visibly
shifted traces (sales append 1.190, returns 1.204). The conflict-aware variant
removes the trace and milder-OOD failures and is best on skew evolution, but it
still trails hard/recent8 on the two strongest OOD families: some pre-drift
observations are coincidentally consistent with the recent reference
($\text{conflict}_j\approx 0$) and survive any $\tau$, whereas the recent-window
age cutoff removes them unconditionally. Reducing $\tau$ to $0.015$ leaves
single-column accuracy unchanged (1.403) and improves batch/range only
marginally (1.322 / 1.232), confirming this is an inherent property of
consistency-based gating rather than a tuning artifact. The calibrated router
sidesteps the issue: it routes to hard/ISOMER on these workloads (the soft
candidate has higher residual), so router safety equals the deployed Hybrid.

\subsection{Downstream Estimators and PostgreSQL Plan-Shape}

On the FactorJoin join kernel, conflict-aware soft alone gives all-drift join
Q-error 1.0367 (recent-window soft 1.038), still behind hard projection 1.024
and the residual-gated router 1.0174, which never selects soft because soft's
join feedback residual (0.0153) exceeds hard projection's (0.0070). On the
six-estimator composition family, conflict-aware soft is within 1--2\% of hard
projection on every estimator (independence 1.222 vs.\ 1.191; IPF/Sinkhorn 1.284
vs.\ 1.280; Gaussian 1.239 vs.\ 1.224; Clayton 1.273 vs.\ 1.256; Gumbel 1.245
vs.\ 1.234; Frank 1.274 vs.\ 1.260), so the router again prefers hard/Hybrid.

\input{../experiments/results/stage2_calibration_20260531/table_stage2_calibration_pg.tex}

Table~\ref{tab:stage2-app-pg} confirms that adding the soft candidate to the
router introduces no PostgreSQL plan-shape regression: the calibrated router is
byte-identical to the Hybrid across all 504 queries. The residual gate's
preference for ISOMER over hard projection on these planner-facing cases (which
is why both the Hybrid and the router show 8.17\% new deviations versus hard
projection's 6.86\%) is a property of the residual criterion, not of the soft
candidate; a plan-shape-aware gate is left to future work.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Detailed Overhead Breakdown}
\label{sec:overhead-app}
%% ═════════════════════════════════════════════════════════════════════════════

\begin{table}[h]
\centering
\caption{OASIS Runtime Overhead Breakdown (1000 trials)}
\label{tab:overhead}
\begin{tabular}{l r r r}
\toprule
Component & p50 (ms) & p95 (ms) & p99 (ms) \\
\midrule
Feature tensorization       & 0.38 & 0.42 & 0.45 \\
Model forward pass          & 0.66 & 0.72 & 0.78 \\
Total correction            & 1.04 & 1.18 & 1.25 \\
Statistics-format conversion & 0.07 & 0.07 & 0.07 \\
Combined correction + conversion & 1.11 & 1.25 & 1.32 \\
\midrule
Feedback collection (per tuple batch) & 0.12 & 0.18 & 0.23 \\
\bottomrule
\end{tabular}
\end{table}

For comparison, a column-level \texttt{ANALYZE} on a $10^7$-row table costs an
estimated 300--500\,ms of sequential I/O. OASIS's combined latency of
$\approx$1.13\,ms is ${\approx}300\times$ cheaper per invocation, well within
the 200\,ms planning budget (desideratum \textbf{D2}).

The statistics-format conversion interface adds negligible overhead
(0.066--0.073\,ms with PostgreSQL-like defaults of 100 MCV entries and 10--20
residual bins). Even in the worst tested configuration (50 MCV, 50 bins), the
conversion completes in 1.03\,ms.

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
