%!TEX encoding = UTF-8 Unicode
\documentclass[runningheads]{llncs}

%% ── packages ─────────────────────────────────────────────────────────────────
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{makecell}
\usepackage{graphicx}
\usepackage{adjustbox}
\usepackage{subcaption}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithm}
\usepackage{algpseudocode}
\let\REQUIRE\Require
\let\ENSURE\Ensure
\let\STATE\State
\let\FOR\For
\let\ENDFOR\EndFor
\let\RETURN\Return
\usepackage{listings}
\usepackage{xcolor}
\usepackage{microtype}
\usepackage{enumitem}
\setlist{nosep,leftmargin=*}
\usepackage{tikz}
\usepackage{pgfplots}
\pgfplotsset{compat=1.18}
\usepackage{hyperref}

% Absorb small residual overfull boxes from long hyphenated compound terms.
\emergencystretch=3em

%% ── reduce spacing ───────────────────────────────────────────────────────────
\setlength{\textfloatsep}{6pt plus 2pt minus 2pt}
\setlength{\floatsep}{4pt plus 2pt minus 2pt}
\setlength{\intextsep}{4pt plus 2pt minus 2pt}
\setlength{\abovecaptionskip}{3pt}
\setlength{\belowcaptionskip}{2pt}
\setlength{\parskip}{2pt plus 1pt minus 1pt}
\setlength{\parsep}{2pt plus 1pt minus 1pt}
\setlength{\itemsep}{1pt plus 1pt minus 1pt}

%% TikZ libraries for figures
\usetikzlibrary{shapes.geometric,arrows.meta,positioning,fit,backgrounds,calc}

%% ── listings style ───────────────────────────────────────────────────────────
\lstset{
  basicstyle=\ttfamily\scriptsize,
  breaklines=true,
  frame=single,
  numbers=left, numberstyle=\tiny,
  keywordstyle=\color{blue!70!black}\bfseries,
  commentstyle=\color{gray},
  stringstyle=\color{teal},
}

\begin{document}

\title{OASIS: Supplementary Materials}

\titlerunning{OASIS: Supplementary Materials}

\author{Author Names Redacted for Review}

\authorrunning{Author Names Redacted}

\institute{Affiliation Redacted\\
\email{redacted@review.org}}

\maketitle

This document collects supplementary materials for the paper
``OASIS: Repairing Stale Optimizer Statistics in the Feedback Nullspace.''
It provides detailed scope and error-budget positioning (\S\,A),
including the marginal-to-joint composition boundary,
model architecture and instrumentation details (\S\,B),
baseline fairness discussion and QuickSel-H adaptation (\S\,C),
the PostgreSQL-style statistics-format adapter and conversion protocol (\S\,D),
predicate-coverage sensitivity analysis (\S\,E),
observation aggregation study details (\S\,F),
PostgreSQL planner-only statistics-injection protocol details (\S\,G),
OOD drift realism protocol (\S\,H),
benchmark-inspired DML trace and public production-telemetry trace protocols
(\S\,I--J), feedback-budget, feedback-noise, and deployment-safety diagnostic
protocols (\S\,K), Stage-2 calibration, estimator-agnostic prior-source swap,
and feedback-locality details (\S\,L), and detailed overhead breakdowns
(\S\,M).

%% ═════════════════════════════════════════════════════════════════════════════
\section{Scope and Error Budget}
\label{sec:scope-app}
%% ═════════════════════════════════════════════════════════════════════════════

OASIS targets \emph{single-column histogram staleness}.
We explicitly exclude from scope:
\begin{itemize}
  \item \textbf{Multi-column correlation errors}: Join cardinality misestimation
        caused by correlated column predicates requires multi-dimensional
        statistics (e.g., mvdists, DNF-based sketches, or learned joint models).
        OASIS operates on one-dimensional marginals and is complementary to
        multi-column methods.
  \item \textbf{Physical design errors}: Missing indexes, suboptimal partitioning,
        or incorrect materialized views are outside the statistics layer.
  \item \textbf{Cost-model calibration}: Even with correct selectivities, the
        optimizer's cost model may mispredict operator runtimes.
        OASIS improves selectivity estimates but does not modify the cost model.
  \item \textbf{Data distribution shifts beyond drift}: Fundamental schema changes
        (column additions, type changes) or massive bulk loads that alter the
        distribution shape beyond what a compound drift operator captures are
        out of scope; these scenarios require a fresh \texttt{ANALYZE}.
\end{itemize}

\paragraph{Error budget decomposition.}
In a production optimizer, the total cardinality estimation error decomposes as:
\[
  \begin{split}
  \text{Total Error} ={}& \underbrace{\text{Single-Column Error}}_{\text{OASIS target}}
  + \underbrace{\text{Correlation Error}}_{\text{multi-column methods}} \\
  &+ \underbrace{\text{Cost-Model Error}}_{\text{calibration}}
  + \underbrace{\text{Physical Design Error}}_{\text{indexing/partitioning}}.
  \end{split}
\]
OASIS addresses the first term. Even perfect single-column statistics do not
eliminate the remaining terms, but accurate marginals are a necessary
prerequisite: learned multi-column estimators (e.g., copula-based models like
CoLSE~\cite{rathuwadu2025colse}) combine one-dimensional marginals with a
learned dependence structure, making single-column statistics quality a
foundational concern.

\paragraph{Marginal-to-joint composition boundary.}
The multi-column composition experiments in the main paper should be interpreted
as downstream use of corrected marginals, not as evidence that OASIS estimates
correlations. The independence estimator uses the maximum-entropy joint
distribution implied by the supplied marginals, while the Gaussian-copula
estimator receives a fixed dependence structure. In this setting, OASIS-noProj
marginals improve the joint estimator only modestly and inconsistently because
they are not always feedback-consistent on the current observation window.
OASIS is therefore the deployable marginal: it starts from the learned
OASIS-noProj marginal, then applies hard feedback-consistency projection before
the marginal is passed to the downstream multi-column estimator. The core
contribution remains single-column statistics correction.

%% ═════════════════════════════════════════════════════════════════════════════
%% Include the existing appendix files
%% ═════════════════════════════════════════════════════════════════════════════

\input{appendix/system_details.tex}

\section{Baseline Fairness Discussion}
\label{sec:baseline-fairness}

All feedback-driven methods compared in the main paper consume the same
observation window of size $K{=}16$ and emit corrected quantiles on the same
$B{=}10$ equi-depth histogram grid. No method receives additional information
or computational budget beyond this shared interface.

\paragraph{STHoles.}
We implement the hierarchical refinement algorithm described
in~\cite{bruno2001stholes}. STHoles builds a tree of bucket refinements from
feedback. We initialize it from the stale prior and allow up to $K$ refinement
steps per correction instance, using the same observation tuples provided to
OASIS. The output quantiles are extracted from the refined tree on the standard
$B{-}1$ grid.

\paragraph{ISOMER.}
We implement the consistency-based histogram adjustment
from~\cite{markl2007consistent}. ISOMER adjusts bucket frequencies to satisfy
observed selectivity constraints while minimizing KL divergence from the prior.
We provide the same $K{=}16$ observation tuples and extract quantiles on the
standard grid.

\paragraph{QuickSel-H.}
See \S\,\ref{sec:system-details-app} for the detailed QuickSel-H adaptation.

\paragraph{Simple heuristics.}
LinInterp and FeedAvg consume the same observation window but use fixed
aggregation rules (blend factor 0.5 and exponential moving average,
respectively). They are included to demonstrate that learned correction is
necessary, not merely that any feedback-driven approach helps.

No method receives per-table or per-distribution tuning; all use default
hyperparameters. Results are averaged over three independent training seeds
(42, 123, 456) with 95\% confidence intervals.

\input{appendix/mcv_adapter.tex}

%% ═════════════════════════════════════════════════════════════════════════════
\section{Predicate-Coverage Sensitivity Analysis}
\label{sec:coverage-sensitivity-app}
%% ═════════════════════════════════════════════════════════════════════════════

OASIS correction quality depends on the observation window covering the value range
relevant to future predicates. The dependence on \emph{how much} of the domain the
feedback constrains is exactly the nullspace effect the main paper characterizes: the
gain concentrates on predicates far from any feedback (the feedback-locality diagnostic,
\S\,\ref{sec:stage2-calibration-app}) and the sparse-feedback regime is where the learned
completion helps most on a real planner (the PostgreSQL budget sweep in the main paper).
Where feedback is too sparse to constrain the queried range, the residual-gated Router
falls back to the maximum-entropy projection, so accuracy degrades gracefully rather than
catastrophically.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Observation Aggregation Study}
\label{sec:aggregation-app}
%% ═════════════════════════════════════════════════════════════════════════════

The imputation layer aggregates the unordered feedback window with a
permutation-invariant set encoder, so that the same model consumes any window size and is
order-independent. The architecture is deliberately secondary to the training objective:
the main paper's ablation (Section on the composite-objective ablation) holds the model and
projection fixed and shows that a reconstruction-only objective is worse than the classical
priors at every feedback budget, while the composite optimizer-facing objective produces
the gain. In other words, the decisive design choice is \emph{what the prior is trained to
optimize}, not the pooling mechanism.

%% ═════════════════════════════════════════════════════════════════════════════
\section{PostgreSQL Planner-Only Statistics-Injection Protocol}
\label{sec:optimizer-protocol-app}
%% ═════════════════════════════════════════════════════════════════════════════

The planner-only evaluation protocol uses PostgreSQL's \texttt{EXPLAIN}
(without \texttt{ANALYZE}) to validate the
statistics$\rightarrow$estimated-cardinality$\rightarrow$plan-shape path.
\texttt{EXPLAIN} invokes PostgreSQL's optimizer and cost model, but it does not
execute the query. The protocol therefore provides optimizer-facing evidence
without making wall-clock runtime claims.

\paragraph{Protocol steps.}
\begin{enumerate}
  \item Create a synthetic PostgreSQL schema with a fact table and dimension
        table. The fact table stores the current post-drift data.
  \item Generate stale and fresh single-column distributions for
        \texttt{fact.x}; build ISOMER, OASIS-noProj, OASIS, and Hybrid corrected
        marginal statistics from the same feedback window.
  \item For each statistics state, create an analyzed one-column source table
        and copy its typed \texttt{pg\_statistic} row into the target
        \texttt{fact.x} statistics entry. This avoids hand-writing
        PostgreSQL's \texttt{anyarray} catalog fields.
  \item For each test query, obtain true cardinality using \texttt{COUNT(*)}
        and obtain PostgreSQL estimated rows and plan shape using
        \texttt{EXPLAIN (FORMAT JSON)}.
  \item Compare row-estimation Q-error, plan-shape match to the
        fresh-statistics planner, recovery of stale/fresh plan disagreements,
        and new deviations from fresh when stale already matched fresh.
\end{enumerate}

\paragraph{Results.}
The formal batch covers 12 configurations: two drift directions
(\texttt{left\_shift} and \texttt{right\_shift}), two fact-table sizes
(100K and 200K rows), and three random seeds. Each configuration evaluates
84 scan/join query templates under each statistics state. No query runtime is
measured.

\begin{itemize}
  \item Stale statistics produce aggregate row Q-error 27.768 and match the
        fresh-statistics plan shape on 56.1\% of queries.
  \item OASIS-noProj reduces row Q-error to 3.703 and recovers 82.2\% of
        stale/fresh plan-shape disagreements, but introduces new plan
        deviations in 2.3\% of cases where stale statistics already matched
        fresh.
  \item OASIS reduces row Q-error to 2.362, matches the fresh-statistics
        plan shape on 95.9\% of queries, recovers 92.1\% of stale/fresh plan
        disagreements, and reduces new plan deviations to 1.1\%.
  \item Across configurations, OASIS has mean row Q-error
        $2.377\pm0.269$, close to ISOMER's $2.390\pm0.286$ and below
        OASIS-noProj's $3.894\pm1.252$.
\end{itemize}

This protocol validates the causal chain
\[
  \text{statistics correction} \rightarrow \text{estimated rows}
  \rightarrow \text{cost re-estimation} \rightarrow \text{plan shape}
\]
inside a real PostgreSQL optimizer, while intentionally avoiding runtime
claims that would depend on executor, cache, and system effects outside the
single-column statistics layer.

Representative plan-shape changes from the batch show the mechanism the planner-only
experiment validates: stale statistics can move estimated rows enough to select a
different root operator or join strategy, while OASIS moves the estimated rows and plan
shape back toward the fresh-statistics planner. These plan changes are not used as runtime
evidence.

%% ═════════════════════════════════════════════════════════════════════════════
\section{OOD Drift Realism Protocol}
\label{sec:ood-drift-app}
%% ═════════════════════════════════════════════════════════════════════════════

The OOD drift realism suite evaluates the same OASIS checkpoint used in the
main synthetic experiments. The checkpoint is trained only on the compound drift
generator; no samples from the OOD families are used for retraining or
hyperparameter tuning. Each OOD case follows the same causal protocol as the
main synthetic suite: capture stale quantiles, interleave drift steps with 16
feedback observations, compute final post-drift quantiles, and evaluate future
selectivity estimates against the final distribution.

The six OOD families are deployment-inspired patterns implemented in
\path{extended_drift_generators.py}: batch loading, range shift, skew
evolution, outlier burst, multimodal drift, and seasonal/mixed drift. The
reported table uses 128 held-out cases per family, $B=10$ histogram buckets,
and 64 future CDF probe points per case. This experiment is not intended to
prove that the simulator is a production DML trace; it tests whether OASIS-noProj
and OASIS remain useful when the drift mechanism differs from the training
generator.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Benchmark-Inspired DML Trace Protocol}
\label{sec:trace-drift-app}
%% ═════════════════════════════════════════════════════════════════════════════

The trace-grounded sanity check is not a production-log experiment. It
constructs explicit DML event streams whose table/column stories are anchored
to analytical benchmark maintenance patterns: append-heavy sales dates,
promotion price revisions, inventory restocking, returns and cancellations,
customer segment churn, and seasonal mixed maintenance. Values are normalized
to the unit interval so that the same single-column correction interface is
used as in the synthetic and OOD experiments.

Each trace case begins with a stale histogram, then executes 16 DML steps. Each
step emits concrete insert, update, and delete counts and then records one
predicate feedback observation against the current post-step table. The
reported run uses 96 cases per trace, $B=10$ histogram buckets, and 64 future
CDF probe points per case. The same OASIS checkpoint trained on compound drift
is used unchanged; no trace-specific retraining or hyperparameter tuning is
performed.

\input{../experiments/results/trace_grounded_drift_v3/table_trace_grounded_drift.tex}

The table should be interpreted as a realism sanity check rather than a claim
that the traces match a particular deployment. Its main role is to expose a
useful boundary: OASIS and Hybrid consistently improve over stale
statistics, while OASIS-noProj can regress on mild update-heavy traces where
feedback consistency is more valuable than learned global extrapolation.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Public Production-Telemetry Trace Protocol}
\label{sec:public-trace-app}
%% ═════════════════════════════════════════════════════════════════════════════

The public production-telemetry case study uses the NASA Kennedy Space Center
HTTP access log distributed by the Internet Traffic Archive~\cite{itaNASAHTTP}.
The trace contains real web-server requests from July 1995 and is a standard
workload-characterization data source~\cite{arlitt1996web}. We use it because
the project does not include private production DBMS workload logs. Accordingly,
the experiment should be read as a trace-grounded append-only analytics-table
case study, not as evidence from a live database deployment or a private query
workload.

The experiment parses the first 750{,}000 valid requests from the July trace
(1995-07-01 00:00:01 to 1995-07-11 13:12:57, UTC-04:00), skipping five malformed
lines. Each HTTP request appends one event. We derive three normalized
single-column signals from real request fields: \emph{reply bytes} is
$\log(1+\text{bytes})$ scaled by the 99.5th percentile, \emph{time of day} is
seconds since midnight divided by 86{,}400, and \emph{path length} is
$\log(1+\text{request-path length})$ scaled by the 99.5th percentile. The
resulting values use the same unit-interval marginal-statistics interface as
the other OASIS experiments.

For each signal we draw 96 windows. A window builds stale statistics from 2{,}000
initial events, then performs 16 append-and-feedback steps with 1{,}000 later
real events per step, for an 800\% append growth relative to the stale prefix.
The windows are stratified across the single public trace and may overlap; we
therefore report this as an external-validity case study, not as an independent
sample for significance testing.
The feedback predicates are sampled from the current post-append table using the
same single-column predicate mix as the other trace experiments. Future
selectivity Q-error is evaluated on 64 predicate constants, half drawn from the
held-out later requests in the same trace window and half from the unit
interval. The trained OASIS checkpoint, hard projection, Soft operator, Hybrid
rule, and Router are used unchanged; there is no
trace-specific retraining or hyperparameter tuning.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Feedback Budget, Feedback Noise, and Safety Diagnostics}
\label{sec:feedback-sensitivity-app}
%% ═════════════════════════════════════════════════════════════════════════════

The feedback-budget experiment truncates each held-out sample to the $K$ most
recent observations, with $K\in\{2,4,8,16\}$. The trained OASIS checkpoint is
not retrained for smaller budgets; instead, the truncated window is padded to
the checkpoint's native 16-slot tensor width. ISOMER, OASIS, and Hybrid
receive the same truncated observation window, so the comparison isolates the
effect of feedback availability.

The feedback-noise experiment perturbs only the observed actual selectivities
in the feedback window. It applies multiplicative log-normal noise with
$\sigma\in\{0,0.02,0.05,0.10\}$, keeps stale priors and fresh future
selectivities fixed, and evaluates the same generator-driven
optimizer-decision proxy used in the main paper. The reported diagnostic uses
64 held-out cases per drift intensity and one noise seed, because noisy
projection constraints are more expensive than the clean-feedback setting.

Both diagnostics are optimizer-facing rather than runtime-facing. The scan and
join decisions are selected from estimated selectivities under a fixed cost
proxy, then scored by true proxy cost using the generated post-drift
distribution. The purpose is to characterize deployment sensitivity, not to
claim wall-clock speedups.

The deployment-safety table in the main paper is derived from cached outputs of
four diagnostics: PostgreSQL planner-only batch summaries, feedback-budget
sensitivity, feedback-noise robustness, and the benchmark-inspired DML trace
sanity check. It does not introduce an oracle selector. Projection uses only the
current feedback window, and Hybrid chooses among stale, ISOMER, OASIS-noProj,
and OASIS by minimizing feedback-window residual. The table is generated by
\path{experiments/deployment_safety_analysis.py}.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Stage-2 Calibration: Soft, Banded Projection, and Router}
\label{sec:stage2-calibration-app}
%% ═════════════════════════════════════════════════════════════════════════════

This section gives the protocol and full per-family results behind the main
paper's finding that the Stage-2 calibration is regime-dependent.

\subsection{Operators}

Stage-2 operators are defined for any supplied marginal prior. In the fixed-prior
comparison below, the supplied prior is the OASIS-noProj histogram; the
prior-source swap later in this section varies that input. All operators act on
the same interval partition induced by the active feedback predicates. \emph{Hard
projection} (the default exact Stage-2 operator) performs cyclic I-projections so
that every active predicate's interval mass matches its observed selectivity
exactly, dropping the oldest constraints when the active set becomes infeasible.
\emph{Soft projection} instead minimizes
$\mathrm{KL}(p\,\|\,p_{\text{prior}}) + \lambda\sum_j w_j (A_j p - y_j)^2$ by
mirror descent over cell masses, where $A_j$ is the interval mask of feedback
$j$ and $y_j$ its observed selectivity. We study three weightings: full-window
($w_j{=}1$), a fixed recent window ($w_j{=}1$ for the last $k{=}8$ observations,
$0$ otherwise), and \emph{conflict-aware}. The conflict-aware weighting fits a
hard reference distribution $p_{\text{ref}}$ to the most recent $k$ observations,
scores each constraint by $\text{conflict}_j = |A_j p_{\text{ref}} - y_j|$, and
sets $w_j = \exp(-(\text{conflict}_j/\tau)^2)$ with $\tau{=}0.03$, so feedback
contradicted by the recent observations is suppressed while old-but-consistent
feedback is retained. \emph{Banded projection} I-projects $p_{\text{prior}}$
onto the bands $y_j\pm\kappa\sqrt{y_j(1-y_j)}$ ($\kappa{=}0$ recovers hard
projection). In the fixed-prior comparison, \emph{Router} selects,
per case, the candidate with the lowest feedback-window residual among
$\{$stale, ISOMER, OASIS-noProj, OASIS, Soft$\}$; in the
prior-source swap, the raw supplied prior replaces OASIS-noProj in the same
protocol. All operators and the diagnostic are implemented in
\path{cdf_kll_ml_pipeline/modern_baselines.py} (\path{correct_soft_isomer},
\path{correct_band_isomer}) and \path{experiments/oasis_soft_projection_diagnostics.py}.

\subsection{Evaluation Protocol}

The single-column comparison reuses the cached compound-drift data and the fixed
OASIS checkpoint; it does not retrain. For each case it generates held-out future
predicates, estimates their selectivity from each Stage-2 output, and reports the
geometric-mean Q-error, together with the feedback residual on the observation
window, quantile MAE against fresh statistics, and an optimizer-proxy join
regret. This future-predicate aggregation differs from the
\texttt{single\_column\_projection} table in the main results (which reports a
different per-case aggregation), so the numbers here are not directly comparable
to that table and are presented as a self-contained operator comparison. The
downstream and safety suites (OOD drift, DML trace, FactorJoin, composition
family, optimizer proxy, PostgreSQL planner subset) reuse the same protocols
documented in the preceding sections, with the soft/router methods added.

\subsection{Single-Column and Optimizer-Proxy Results}

\input{../experiments/results/optimizer_decision_proxy_v3/table_optimizer_decision_proxy.tex}

Table~\ref{tab:optimizer_decision_proxy} reports the operator comparison on the
generator-driven optimizer-decision proxy. The ordering is
stale ($2.867$) $\to$ ISOMER ($1.664$) $\to$ hard projection of the learned prior
($1.401$) $\to$ Soft ($1.362$) $\to$ Router ($1.375$, $92.9\%$ join-optimal),
approaching fresh. Soft has the lowest single-column Q-error but, as the main-paper
ablation shows, the gain over hard projection is small and is not the source of OASIS's
advantage; the Router recovers most of Soft's accuracy while remaining deployment-safe and
never selecting a higher-residual candidate. As a negative control, banded projection
(relaxing the equality constraints to confidence bands) monotonically \emph{worsens}
single-column Q-error as the band widens, confirming that the gain comes from per-case
routing rather than from relaxing the feedback constraints.

\subsection{Estimator-Agnostic Prior-Source Swap}

The main text also asks whether Stage 2 is merely an OASIS-MLP-specific patch.
The prior-source swap uses the same cached drift cases and the same held-out
future-predicate protocol as the Stage-2 comparison, but replaces the supplied
marginal prior with six alternatives: stale statistics, LinInterp, FeedAvg,
STHoles, QuickSel-H, and the OASIS MLP. Each source is evaluated raw, after hard
projection initialized from that source, and after Router over
$\{$stale, ISOMER, raw prior, hard projection, Soft$\}$.

\input{../experiments/results/stage1_estimator_swap_v3/table_stage1_estimator_swap.tex}

Table~\ref{tab:stage1_estimator_swap} shows that the calibration layer is
estimator-agnostic in aggregate. Hard projection improves every raw prior, and
Router further improves every source-level row, reducing future Q-error by
$10.8\%$--$41.9\%$ relative to the corresponding raw marginal, with the final Router
outputs clustered tightly ($1.39$--$1.44$). The OASIS prior is the strongest \emph{raw}
input ($1.489$, below STHoles $1.572$, QuickSel-H $1.614$, and the recent learned
query-driven baseline LQM $1.607$) and routes to the best final value ($1.328$, versus
$1.425$ for LQM). Fitting a flexible learned model to the feedback alone (LQM) therefore
lands with the classical self-tuning histograms; the advantage comes from training the
completion against optimizer error, while the reusable mechanism is the calibration layer
that accepts several imperfect marginal sources.

\subsection{Feedback-Locality Diagnostic}

To see where the learned imputation helps after projection, we bin the
held-out future predicates by directed Hausdorff distance from their endpoints to
the nearest feedback endpoint in fresh-CDF space. This is the spatial signature of a
nullspace completion: OASIS beats ISOMER in every locality bin, and the advantage
\emph{grows} with distance from feedback---$+9.1\%$, $+14.5\%$, and $+21.2\%$ in the
near, mid, and far bins. The constraints determine the marginal near the observations,
and the learned prior earns its advantage where they do not.

\input{../experiments/results/proj_v3/table_feedback_locality.tex}

\subsection{Feedback Budget and Noise}

The feedback-budget and feedback-noise diagnostics (protocols above) run on the
optimizer-decision proxy. Table~\ref{tab:feedback_budget_sensitivity} sweeps the visible
window $K$: the learned completion beats ISOMER at every budget.
Table~\ref{tab:feedback_noise_robustness} perturbs the observed selectivities; at $10\%$
multiplicative noise the Router degrades gracefully (selectivity Q-error
$1.375\to1.418$, join-optimal $92.9\%\to92.2\%$), staying well below ISOMER ($1.758$) and
stale ($2.867$).

\input{../experiments/results/feedback_budget_sensitivity_v3/table_feedback_budget_sensitivity.tex}

\input{../experiments/results/feedback_noise_robustness_v3/table_feedback_noise_robustness.tex}

\subsection{Composition, Joins, and Plan-Shape Safety}

The downstream-safety results are reported in full in the main paper: the six-estimator
composition family, the FactorJoin join kernel, and the 12-configuration PostgreSQL
plan-shape batch. Across all of these the calibrated marginal (OASIS, Router) is safe,
whereas the raw prior (OASIS-noProj) can regress---it is actively harmful in the bilinear
FactorJoin kernel and on the append-only NASA trace, which is why the projection and
residual gating are part of the method. Adding the Soft candidate to the Router pool
introduces no plan-shape regression: on the dense planner configurations the Router selects
ISOMER and matches the deployed Hybrid, with ISOMER, OASIS, and the Router all near the
same $\approx 1\%$ new-deviation rate. A plan-shape-aware gate, which could prefer hard
projection over ISOMER on planner-facing cases, is left to future work.

%% ═════════════════════════════════════════════════════════════════════════════
\section{Detailed Overhead Breakdown}
\label{sec:overhead-app}
%% ═════════════════════════════════════════════════════════════════════════════

\begin{table}[h]
\centering
\caption{OASIS Runtime Overhead Breakdown (1000 trials)}
\label{tab:overhead}
\begin{tabular}{l r r r}
\toprule
Component & p50 (ms) & p95 (ms) & p99 (ms) \\
\midrule
Feature tensorization       & 0.38 & 0.42 & 0.45 \\
Model forward pass          & 0.66 & 0.72 & 0.78 \\
Total correction            & 1.04 & 1.18 & 1.25 \\
Statistics-format conversion & 0.07 & 0.07 & 0.07 \\
Combined correction + conversion & 1.11 & 1.25 & 1.32 \\
\midrule
Feedback collection (per tuple batch) & 0.12 & 0.18 & 0.23 \\
\bottomrule
\end{tabular}
\end{table}

For comparison, a column-level \texttt{ANALYZE} on a $10^7$-row table costs an
estimated 300--500\,ms of sequential I/O. OASIS's combined latency of
$\approx$1.13\,ms is ${\approx}300\times$ cheaper per invocation, well within
the 200\,ms planning budget (desideratum \textbf{D2}).

The statistics-format conversion interface adds negligible overhead
(0.066--0.073\,ms with PostgreSQL-like defaults of 100 MCV entries and 10--20
residual bins). Even in the worst tested configuration (50 MCV, 50 bins), the
conversion completes in 1.03\,ms.

\bibliographystyle{splncs04}
\bibliography{references}

\end{document}
